---
title: "Self-restraint as mediator - Anxiety"
output: html_document
date: "`r Sys.Date()`
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(semPlot)
library(lavaan)
library(psych)
library(gtsummary)
```

#1. Cleaning
```{r}
df <- read.csv("lgbtPKU.csv")
dim(df)
```

```{r}
#renaming for categoricals and ordinals
df <- df %>%
  rename(
         grade = Q7,
         sexual_orientation = Q12,
         nonconformity_masculine = Q14,
         nonconformity_feminine = Q15,
         support_from_queer = Q25_行1,
         support_from_nonqueer = Q25_行1,
         )

#manipulations
df$age <- 33 - df$Q3
df$mainland <- ifelse(df$Q4 == 1, 1, 0)
df$han_ethnicity <- 2- df$Q5
df$department <- factor(df$Q6)
df$urban_residency <- ifelse(df$Q8 == 1, 1, 0)
df$self_restraint_teacher <- 5 - df$Q18
df$documented_male <- 2 - df$Q9
df$documented_female <- 1 - df$documented_male

df$assigned_male <- ifelse(df$Q10 != 3, 2 - df$Q10, NA)
df$assigned_female <- 1 - df$assigned_male

df$cisgender <- 2 - df$Q11
df$transgender <- 1 - df$cisgender

library(fastDummies)
df <- dummy_cols(df, select_columns = "sexual_orientation", remove_first_dummy = TRUE)

df$singleness <- ifelse(df$Q13 == 1, 1, 0)

df$gn <- ifelse(df$nonconformity_masculine == -3, 
             df$nonconformity_feminine, df$nonconformity_masculine)

df$outness <- 2 - df$Q16
df$out_to_teacher <- ifelse(df$Q17 == 1, 1, 0)
df$public_out = ifelse(df$Q21 == 1, 1, 0)
df$out_to_queer_friends = ifelse(df$Q23 == -3, 1, df$Q23)
df$out_to_nonqueer_friends = ifelse(df$Q24 == -3, 1, df$Q24)

#LGBTQ+
df$LGBTQ <- ifelse(df$Q10 != 3 & df$cisgender == 1 & df$sexual_orientation == 2, 0, 1)

#perceived LGBTQ+-affirming attitude
df$perceived_acceptance_teacher = 5 - df$Q19
df$perceived_acceptance_student = 5 - df$Q28

df$victimization_teacher <- 1 - df$Q20_选项17
df$victimization_student <- 1 - df$Q29_选项14
df$victimization_ever <- ifelse(df$victimization_teacher == 1 | df$victimization_student == 1, 1, 0)

#Self-restraint of gender expression
df$self_restraint_allstudent <- 10 - df$Q27_行1 - df$Q27_行2
df$self_restraint_otherLGBTQ <- 5 - df$Q27_行1
df$self_restraint_otherstudents <- 5 - df$Q27_行2
df$self_restraint <- df$self_restraint_teacher + df$self_restraint_allstudent - 3

# self-esteem: Rosenberg Self-Esteem Scale, RSES
df <- df %>%
  mutate(
    Q31_行3_r  = 6 - Q31_行3,
    Q31_行5_r  = 6 - Q31_行5,
    Q31_行8_r  = 6 - Q31_行8,
    Q31_行9_r  = 6 - Q31_行9,
    Q31_行10_r = 6 - Q31_行10)

df <- df %>%
  mutate(
    self_esteem = rowSums(
      select(., all_of(c(
        "Q31_行1", "Q31_行2", "Q31_行3_r", "Q31_行4", "Q31_行5_r",
        "Q31_行6", "Q31_行7", "Q31_行8_r", "Q31_行9_r", "Q31_行10_r"
      ))),
      na.rm = TRUE
    )
  )

# depression: Center for Epidemiologic Studies Depression Scale, CESD 9-item Chinese version
df <- df %>%
  mutate(
    Q30_行5_r = 5 - Q30_行5,
    Q30_行7_r = 5 - Q30_行7)

df <- df %>%
  mutate(
    depression = Q30_行1 + Q30_行2 + Q30_行3 + Q30_行4 + 
                 Q30_行5_r + Q30_行6 + Q30_行7_r + Q30_行8 + Q30_行9
  )

#anxiety: Generalized Anxiety Disorder Scale, GAD-7
df <- df %>%
  mutate(
    anxiety = rowSums(select(., all_of(c("Q32_行1", "Q32_行2", "Q32_行3", "Q32_行4", "Q32_行5", "Q32_行6", "Q32_行7"))),
                      na.rm = TRUE))
```


```{r}
#check inconsistent cases
#transgender
df %>%
  filter((documented_female == 1 & assigned_male == 1 & transgender == 0) | (documented_male == 1 & assigned_female == 1 & transgender == 0))

#outness
df %>%
  filter((outness == 0 & out_to_teacher== 1) | (outness == 0 & public_out == 1))


#outness in nonLGBTQ
#SOGIE-based victimization among non-LGBTQ+
df %>%
  filter(LGBTQ == 0 & outness == 1)

df %>%
  filter(LGBTQ == 0 & out_to_teacher == 1)
```

```{r}
#drop one inconsistent case
df1 <- df %>%
  filter(!((documented_female == 1 & assigned_male == 1 & transgender == 0) | (documented_male == 1 & assigned_female == 1 & transgender == 0)) %in% TRUE)
dim(df1)
```
```{r}
#drop two cases that non-LGBTQ+ reported coming out
df1 <- df1 %>%
  filter(!(LGBTQ == 0 & outness == 1))
dim(df1)
```
```{r}
#drop inconsistent case
#constantly filling the same value for multiple question matrices
df1 <- df1 %>%
  filter(!(index == 1273))
dim(df1)
```
```{r}
#drop age under 18 (5 cases)
df1 <- df1 %>%
  filter(!(age < 18))
dim(df1)
```

#2.3. victm
```{r}
#count
df1$victim_teacher_count <- rowSums(df1[, paste0('Q20_选项', 1:15)])
df1$victim_student_count <- rowSums(df1[, paste0('Q29_选项', 1:12)])
df1$victimization_count <- df1$victim_teacher_count + df1$victim_student_count

table(df1$victim_teacher_count)
table(df1$victim_student_count)
table(df1$victimization_count)
```

```{r}
#Victimization weighted sum score
# microaggression *1; moderate violence & discrimination *2; severe violence *3
df1$victm_teacher_sum <- 
  1 * df1$Q20_选项1 + 1 * df1$Q20_选项3 + 1 * df1$Q20_选项4 +
  2 * df1$Q20_选项2 + 2 * df1$Q20_选项6 + 2 * df1$Q20_选项7 + 2 * df1$Q20_选项8 +
  2 * df1$Q20_选项9 + 2 * df1$Q20_选项10 + 2 * df1$Q20_选项11 + 2 * df1$Q20_选项12 + 2 * df1$Q20_选项15 +
  3 * df1$Q20_选项5 + 3 * df1$Q20_选项13 + 3 * df1$Q20_选项14
  
df1$victm_student_sum <-
  1 * df1$Q29_选项1 + 1 * df1$Q29_选项3 + 1 * df1$Q29_选项4 +
  2 * df1$Q29_选项2 + 2 * df1$Q29_选项6 + 2 * df1$Q29_选项7 + 2 * df1$Q29_选项8 +
  2 * df1$Q29_选项9 + 2 * df1$Q29_选项12 +
  3 * df1$Q29_选项5 + 3 * df1$Q29_选项10 + 3 * df1$Q29_选项11

df1$victimization_sum <- df1$victm_teacher_sum + df1$victm_student_sum

table(df1$victm_teacher_sum)
table(df1$victm_student_sum)
table(df1$victimization_sum)

df1 %>% slice_max(victm_teacher_sum, n = 10)
df1 %>% slice_max(victm_student_sum, n = 10)
df1 %>% slice_max(victimization_sum, n = 10)
```
```{r}
describe(df1$victimization_sum)
aggregate(victimization_sum ~ LGBTQ, data =df1, mean)
aggregate(victimization_sum ~ LGBTQ, data =df1, sd)
```

#4.2 Regression - Anxiety
```{r}
vars_regression <- c("LGBTQ", "documented_female", "victimization_sum", "gn", "self_restraint",
                     "age", "mainland", "han_ethnicity", "department","urban_residency"
                     )
```


`anx_mdl_1`: Anxiety ~ Gender Non-confirmative
```{r}
anx_mdl_1 <- lm(anxiety ~ gn, data = df1)
summary(anx_mdl_1)
```

`anx_mdl_2`: Anxiety ~ Self-Restraint
```{r}
anx_mdl_2 <- lm(anxiety ~ self_restraint, data = df1)
summary(anx_mdl_2)
```

`anx_mdl_3`: Anxiety ~ Self-Restraint + Gender Nonconformity
```{r}
anx_mdl_3 <- lm(anxiety ~ self_restraint + gn, data = df1)
summary(anx_mdl_3)
```

`anx_mdl_4`: Anxiety ~ Self-Restraint + Gender Nonconformity + victimization_sum + documented_female + LGBTQ
```{r}
anx_mdl_4 <- lm(anxiety ~ self_restraint + gn + victimization_sum + documented_female + LGBTQ, data = df1)
summary(anx_mdl_4)
```

`anx_mdl_5`: Anxiety ~ Self-Restraint + Gender Nonconformity + victimization_sum + documented_female + LGBTQ + self_restraint*documented_female 
```{r}
anx_mdl_5 <- lm(anxiety ~ self_restraint + gn + victimization_sum + documented_female + LGBTQ + self_restraint*documented_female, data = df1)
summary(anx_mdl_5)
```

# 6. S2 GN
```{r}
#simple linear regression: Mental health ~ Gender nonconformity
s2s1_dep_model <- lm(depression ~ gn, data = df1)
summary(s2s1_dep_model)

s2s1_anx_model <- lm(anxiety ~ gn, data = df1)
summary(s2s1_anx_model)

summary(lm(depression ~ documented_female, data=df1))
summary(lm(victimization_sum ~ documented_female, data=df1))
summary(lm(self_restraint ~ documented_female, data=df1))

```
## 6._.1. no interaction term

### 6.1.1.0 S2 dep (doc_female controlled for all var)
```{r}
s2_dep_6.1.1.0_model <- '
  depression ~ a*gn + b*victimization_sum + c*self_restraint + documented_female
  
  victimization_sum ~ d*gn + documented_female
  self_restraint ~ e*gn + documented_female

  gn ~ documented_female
  
  #effects
  indirect_victimization := d*b
  indirect_restraint := e*c
  total_indirect := b*d + c*e
  total_effect := a + total_indirect
'

s2_dep_6.1.1.0_fit <- sem(s2_dep_6.1.1.0_model, data=df1, estimator="ML", se="bootstrap" , bootstrap=1000)
```

```{r}
summary(s2_dep_6.1.1.0_fit, standardized=TRUE, fit.measures=TRUE)
```

```{r}
semPaths(s2_dep_6.1.1.0_fit, intercept = FALSE, what = "std", whatLabels="std", reorder = FALSE, 
         rainbow = .5, rotation = 2, layout = "spring")
```

### 6.1.1.5 S2 dep (doc_female controlled for one var)
```{r}
s2_dep_6.1.1.5_model <- '
  depression ~ a*gn + b*victimization_sum + c*self_restraint
  
  victimization_sum ~ d*gn
  self_restraint ~ e*gn + documented_female
  
  #effects
  indirect_victimization := d*b
  indirect_restraint := e*c
  total_indirect := b*d + c*e
  total_effect := a + total_indirect
'

s2_dep_6.1.1.5_fit <- sem(s2_dep_6.1.1.5_model, data=df1, estimator="ML", se="bootstrap" , bootstrap=1000)
```

```{r}
summary(s2_dep_6.1.1.5_fit, standardized=TRUE, fit.measures=TRUE)
```

```{r}
semPaths(s2_dep_6.1.1.5_fit, intercept = FALSE, what = "std", whatLabels="std", reorder = FALSE, 
         rainbow = .5, rotation = 2, layout = "spring")
```

### 6.2.1.0 S2 anx (doc_female controlled for all var)
```{r}
s2_anx_6.2.1.0_model <- '
  anxiety ~ a*gn + b*victimization_sum + c*self_restraint + documented_female
  
  victimization_sum ~ d*gn + documented_female
  self_restraint ~ e*gn + documented_female

  gn ~ documented_female
  
  #effects
  indirect_victimization := d*b
  indirect_restraint := e*c
  total_indirect := b*d + c*e
  total_effect := a + total_indirect
'

s2_anx_6.2.1.0_fit <- sem(s2_anx_6.2.1.0_model, data=df1, estimator="ML", se="bootstrap" , bootstrap=1000)
```

```{r}
summary(s2_anx_6.2.1.0_fit, standardized=TRUE, fit.measures=TRUE)
```

```{r}
semPaths(s2_anx_6.2.1.0_fit, intercept = FALSE, what = "std", whatLabels="std", reorder = FALSE, 
         rainbow = .5, rotation = 2, layout = "spring")
```

### 6.2.1.5 S2 anx (doc_female controlled for only one var)
```{r}
s2_anx_6.2.1.5_model <- '
  anxiety ~ a*gn + b*victimization_sum + c*self_restraint
  
  victimization_sum ~ d*gn
  self_restraint ~ e*gn + documented_female
  
  #effects
  indirect_victimization := d*b
  indirect_restraint := e*c
  total_indirect := b*d + c*e
  total_effect := a + total_indirect
'

s2_anx_6.2.1.5_fit <- sem(s2_anx_6.2.1.5_model, data=df1, estimator="ML", se="bootstrap" , bootstrap=1000)
```

```{r}
summary(s2_anx_6.2.1.5_fit, standardized=TRUE, fit.measures=TRUE)
```

```{r}
semPaths(s2_anx_6.2.1.5_fit, intercept = FALSE, what = "std", whatLabels="std", reorder = FALSE, 
         rainbow = .5, rotation = 2, layout = "spring")
```


## 6._.2 moderation gn*LGBTQ
```{r}
df1$interaction_gn_LGBTQ <- df1$gn * df1$LGBTQ
```

### 6.1.2.0 dep ~ moderation gn*LGBTQ (doc_female cntrl for all)
```{r}
s2_dep_6.1.2.0_model <- '
  depression ~ a*gn + b*victimization_sum + c*self_restraint + documented_female
  
  victimization_sum ~ d*gn + documented_female
  self_restraint ~ e*gn + g*interaction_gn_LGBTQ + documented_female + LGBTQ

  gn ~ documented_female
  
  # indirect effects
  ind_victimization := b*d
  ind_SR_nonLGBTQ := e*c
  ind_SR_LGBTQ := (e+g)*c
  total_ind_nonLGBTQ := b*d + e*c
  total_ind_LGBTQ := b*d + (e+g)*c
  total_nonLGBTQ := a + total_ind_nonLGBTQ
  total_LGBTQ := a + total_ind_LGBTQ
  index_modmed := g*c
'

s2_dep_6.1.2.0_fit <- sem(s2_dep_6.1.2.0_model, data=df1, estimator="ML", se="bootstrap" , bootstrap=1000)
```

```{r}
summary(s2_dep_6.1.2.0_fit, standardized=TRUE, fit.measures=TRUE)
```

```{r}
semPaths(s2_dep_6.1.2.0_fit, intercept = FALSE, what = "std", whatLabels="std", reorder = FALSE, 
         rainbow = .5, rotation = 2, layout = "spring")
```

### 6.1.2.5 dep ~ moderation gn*LGBTQ (doc_female cntrl for one)
```{r}
s2_dep_6.1.2.5_model <- '
  depression ~ a*gn + b*victimization_sum + c*self_restraint
  
  victimization_sum ~ d*gn + documented_female
  self_restraint ~ e*gn + g*interaction_gn_LGBTQ + documented_female + LGBTQ
  
  # indirect effects
  ind_victimization := b*d
  ind_SR_nonLGBTQ := e*c
  ind_SR_LGBTQ := (e+g)*c
  total_ind_nonLGBTQ := b*d + e*c
  total_ind_LGBTQ := b*d + (e+g)*c
  total_nonLGBTQ := a + total_ind_nonLGBTQ
  total_LGBTQ := a + total_ind_LGBTQ
  index_modmed := g*c
'

s2_dep_6.1.2.5_fit <- sem(s2_dep_6.1.2.5_model, data=df1, estimator="ML", se="bootstrap" , bootstrap=1000)
```

```{r}
summary(s2_dep_6.1.2.5_fit, standardized=TRUE, fit.measures=TRUE)
```

```{r}
semPaths(s2_dep_6.1.2.5_fit, intercept = FALSE, what = "std", whatLabels="std", reorder = FALSE, 
         rainbow = .5, rotation = 2, layout = "spring")
```


### 6.2.2.0 anx ~ moderation gn*LGBTQ (doc_female cntrl for all)
```{r}
s2_anx_6.2.2.0_model <- '
  anxiety ~ a*gn + b*victimization_sum + c*self_restraint + documented_female
  
  victimization_sum ~ d*gn + documented_female
  self_restraint ~ e*gn + g*interaction_gn_LGBTQ + documented_female + LGBTQ

  gn ~ documented_female
  
  # indirect effects
  ind_victimization := b*d
  ind_SR_nonLGBTQ := e*c
  ind_SR_LGBTQ := (e+g)*c
  total_ind_nonLGBTQ := b*d + e*c
  total_ind_LGBTQ := b*d + (e+g)*c
  total_nonLGBTQ := a + total_ind_nonLGBTQ
  total_LGBTQ := a + total_ind_LGBTQ
  index_modmed := g*c
'

s2_anx_6.2.2.0_fit <- sem(s2_dep_6.2.2.0_model, data=df1, estimator="ML", se="bootstrap" , bootstrap=1000)
```

```{r}
summary(s2_anx_6.2.2.0_fit, standardized=TRUE, fit.measures=TRUE)
```

```{r}
semPaths(s2_anx_6.2.2.0_fit, intercept = FALSE, what = "std", whatLabels="std", reorder = FALSE, 
         rainbow = .5, rotation = 2, layout = "spring")
```

### 6.2.2.5 anx ~ moderation gn*LGBTQ (doc_female cntrl for one)
```{r}
s2_anx_6.2.2.5_model <- '
  anxiety ~ a*gn + b*victimization_sum + c*self_restraint
  
  victimization_sum ~ d*gn
  self_restraint ~ e*gn + g*interaction_gn_LGBTQ + documented_female + LGBTQ
  
  # indirect effects
  ind_victimization := b*d
  ind_SR_nonLGBTQ := e*c
  ind_SR_LGBTQ := (e+g)*c
  total_ind_nonLGBTQ := b*d + e*c
  total_ind_LGBTQ := b*d + (e+g)*c
  total_nonLGBTQ := a + total_ind_nonLGBTQ
  total_LGBTQ := a + total_ind_LGBTQ
  index_modmed := g*c
'

s2_anx_6.2.2.5_fit <- sem(s2_anx_6.2.2.5_model, data=df1, estimator="ML", se="bootstrap" , bootstrap=1000)
```

```{r}
summary(s2_anx_6.2.2.5_fit, standardized=TRUE, fit.measures=TRUE)
```

```{r}
semPaths(s2_anx_6.2.2.5_fit, intercept = FALSE, what = "std", whatLabels="std", reorder = FALSE, 
         rainbow = .5, rotation = 2, layout = "spring")
```

## 6._.3 Grouped by LGBTQ


```{r}
s2_dep_model <- '
  depression ~ gn + victimization_sum + self_restraint + documented_female
  
  victimization_sum ~ gn + documented_female
  self_restraint ~ gn + documented_female

  gn ~ documented_female
'

s2_dep_fit <- sem(s2_dep_model, data=df1, group = "LGBTQ", estimator="ML", se="bootstrap" , bootstrap=1000)
summary(s2_dep_fit, standardized=TRUE, fit.measures=TRUE)
```

```{r}
s2_anx_model <- '
  anxiety ~ gn + victimization_sum + self_restraint + documented_female
  
  victimization_sum ~ gn + documented_female
  self_restraint ~ gn + documented_female

  gn ~ documented_female
'

s2_anx_fit <- sem(s2_anx_model, data=df1, group = "LGBTQ", estimator="ML", se="bootstrap" , bootstrap=1000)
summary(s2_dep_fit, standardized=TRUE, fit.measures=TRUE)
```

```{r}
lm(self_restraint ~ gn + documented_female,
   data=subset(df1, LGBTQ==0))

lm(self_restraint ~ gn + documented_female,
   data=subset(df1, LGBTQ==1))
```
