---
title: "Self-restraint as mediator - EFA Victimization"
output:
  html_document:
    code_folding: hide
date: "`r Sys.Date()`"
---

# 0. Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(semPlot)
library(lavaan)
library(psych)
library(gtsummary)
library(GPArotation)
library(ggplot2)
library(purrr)
library(ggcorrplot)

```

#1. Cleaning
```{r}

df_r <- read.csv("lgbtPKU.csv")
```

Raw data dim: `r dim(df_r)`

## renaming for categoricals and ordinals

```{r}
df <- df_r |>
  rename(
         grade = Q7,
         sexual_orientation = Q12,
         nonconformity_masculine = Q14,
         nonconformity_feminine = Q15,
         support_from_queer = Q25_行1,
         support_from_nonqueer = Q25_行1
         ) |>
  mutate(
# Sex
    assigned_male = ifelse(Q10 != 3, 2 - Q10, NA),
    assigned_female = 1 - assigned_male,
    documented_male = 2 - Q9,
    documented_female = 1 - documented_male,

# Age
    age = 33 - Q3,

# LGBTQ+
    cisgender = 2 - Q11,
    transgender = 1 - cisgender,
    LGBTQ = ifelse(Q10 != 3 & cisgender == 1 & sexual_orientation == 2, 0, 1),

# outness
    outness = 2 - Q16,
    out_to_teacher = ifelse(Q17 == 1, 1, 0),
    public_out = ifelse(Q21 == 1, 1, 0),
    out_to_queer_friends = ifelse(Q23 == -3, 1, Q23),
    out_to_nonqueer_friends = ifelse(Q24 == -3, 1, Q24),

# Victimization and perceived LGBTQ+ affirming attitude
    perceived_acceptance_teacher = 5 - Q19,
    perceived_acceptance_student = 5 - Q28,

    victimization_teacher = 1 - Q20_选项17,
    victimization_student = 1 - Q29_选项14,
    victimization_ever = ifelse(victimization_teacher == 1 | victimization_student == 1, 1, 0),

#Self-restraint of gender expression
    self_restraint_teacher = 5 - Q18,
    self_restraint_allstudent = 10 - Q27_行1 - Q27_行2,
    self_restraint_otherLGBTQ = 5 - Q27_行1,
    self_restraint_otherstudents = 5 - Q27_行2,
    self_restraint = self_restraint_teacher + self_restraint_allstudent - 3,

# self-esteem: Rosenberg Self-Esteem Scale, RSES
    Q31_行3_r  = 6 - Q31_行3,
    Q31_行5_r  = 6 - Q31_行5,
    Q31_行8_r  = 6 - Q31_行8,
    Q31_行9_r  = 6 - Q31_行9,
    Q31_行10_r = 6 - Q31_行10,

    self_esteem = rowSums(
      pick(all_of(c(
        "Q31_行1", "Q31_行2", "Q31_行3_r", "Q31_行4", "Q31_行5_r", "Q31_行6", "Q31_行7", "Q31_行8_r", "Q31_行9_r", "Q31_行10_r"
      ))),
      na.rm = TRUE
    ),


# depression: Center for Epidemiologic Studies Depression Scale, CESD 9-item Chinese version
    Q30_行5_r = 5 - Q30_行5,
    Q30_行7_r = 5 - Q30_行7,
    depression = Q30_行1 + Q30_行2 + Q30_行3 + Q30_行4 + Q30_行5_r + Q30_行6 + Q30_行7_r + Q30_行8 + Q30_行9,
  
#anxiety: Generalized Anxiety Disorder Scale, GAD-7
    anxiety = rowSums(
      pick(all_of(c("Q32_行1", "Q32_行2", "Q32_行3", "Q32_行4", "Q32_行5", "Q32_行6", "Q32_行7"))), na.rm = TRUE
    )
  )
```

Renamed data dim: `r dim(df)`


## drop inconsistent cases
```{r}

#drop one inconsistent case
df1 <- df |>
  filter(!((documented_female == 1 & assigned_male == 1 & transgender == 0) | (documented_male == 1 & assigned_female == 1 & transgender == 0)) %in% TRUE) |>

#drop two cases that non-LGBTQ+ reported coming out
  filter(!(LGBTQ == 0 & outness == 1)) |> 

#constantly filling the same value for multiple question matrices
  filter(!(index == 1273)) |>

#drop age under 18 (5 cases)
  filter(!(age < 18))
```

Dim after some obs dropped: `r dim(df1)`

#2.3. victm EFA

```{r}
vic_q <- df1 |>
  select(
    # Q20
    Q20_选项1, Q20_选项2, Q20_选项3, Q20_选项4, Q20_选项5,
    Q20_选项6, Q20_选项7, Q20_选项8, Q20_选项9, Q20_选项10,
    Q20_选项11, Q20_选项12, Q20_选项13, Q20_选项14, Q20_选项15,

    # Q29
    Q29_选项1, Q29_选项2, Q29_选项3, Q29_选项4, Q29_选项5,
    Q29_选项6, Q29_选项7, Q29_选项8, Q29_选项9, Q29_选项10,
    Q29_选项11, Q29_选项12
  )

# Give items shorter names for easier interpretation
names(vic_q) <- c(
  paste0("Q20_", 1:15),
  paste0("Q29_", 1:12)
)
```

Questions on Victimization: `r names(vic_q)`

## 1. Check
```{r}
# Find variables that contain only 0, 1, and possibly NA
binary_vars <- vic_q |>
  select(where(~ all(na.omit(.) %in% c(0, 1)))) |>
  names()

# Calculate percentage of 0 and 1 for each variable
plot_data <- vic_q |>
  select(all_of(binary_vars)) |>
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  ) |>
  filter(!is.na(value)) |>
  count(variable, value) |>
  group_by(variable) |>
  mutate(
    percentage = n / sum(n) * 100
  ) |>
  ungroup()

# Plot
plot_data |>

  ggplot(
    aes(x = variable, y = percentage, fill = factor(value))
  ) +
  geom_col(
    width = 0.7
  ) +
  geom_text(
    aes(label = paste0(round(percentage, 1), "%")),
    position = position_dodge(width = 0.8),
    vjust = -0.3,
    size = 3
  ) +
  scale_y_continuous(
    limits = c(0, 105),
    breaks = seq(0, 100, 20),
    labels = function(x) paste0(x, "%")
  ) +
  labs(
    x = NULL,
    y = "Percentage",
    fill = "Value",
    title = "Distribution of 0/1 in Victimization Related Questions"
  ) +
  theme_classic() +
    theme(
      axis.text.x = element_text(
        angle = 45,
        hjust = 1
      )
    ) +
  coord_flip()
```

Slightly too many 0s.
"The problem is that for binary variables, extremely rare 1s provide very little information about how items relate to each other. This becomes especially important in your analysis because you’re using tetrachoric correlations." - chatgpt

## 2. Delete those with 0 variance and with 99% 0 
```{r}

vars_to_remove = names(vic_q)[sapply(vic_q, function(x) {
  x <- x[!is.na(x)]
  length(x) == 0 ||
    length(unique(x)) == 1 ||     # zero variance
    mean(x == 0) >= 0.99         # 99.5% or more are 0
})]

vic_clean = vic_q |>
  select(-all_of(vars_to_remove))
```


Deleted variables: `r vars_to_remove`


## 3. Tetrachoric correlation matrix

For binary items, this is preferable to ordinary Pearson correlations:
```{r}
corr <- tetrachoric(vic_clean)
eigen(corr$rho)$values
ggcorrplot(corr$rho, lab = TRUE)
```

## 4. Scree Plot
```{r}
eigen = fa.parallel(vic_clean, cor='poly', fa = "pc")
```

Eigenvalues
```{r}
round(eigen$pc.values, digit = 2)
```

## 5. EFA

### 3-factor
```{r}
efa_3 <- fa(
  corr$rho,
  nfactors = 3,
  n.obs = nrow(vic_clean),
  fm = "minres",
  rotate = "oblimin"
)
efa_3
efa_3$loadings
```

EFA3: communality and uniqueness 
```{r}
efa_3$communality
efa_3$uniquenesses
```

Communality is the proportion of an item’s variance explained by the common factors in your EFA.
Uniqueness is the portion not explained by the common factors:

***Q20——选项4 Q20_选项15 require further investigation.***

Q20_4   $h^2$ = 1.026 $\leftarrow$ improper
Q20_15  $h^2$ = 1.062 $\leftarrow$ improper
Q29_4   $h^2$ = 0.945 $\leftarrow$ technically proper, but unusually high

```{r}
colMeans(vic_clean[, c("Q20_4", "Q20_15", "Q29_4")], na.rm = TRUE)
round(corr$rho[c("Q20_4", "Q20_15", "Q29_4"), ], 3)
```


### neighbors
```{r}
efa_2 <- fa(
  corr$rho,
  nfactors = 2,
  n.obs = nrow(vic_clean),
  fm = "minres",
  rotate = "oblimin"
)
efa_2
efa_2$loadings
```

EFA2: communality and uniqueness 
```{r}
efa_2$communality
efa_2$uniquenesses
```

```{r}
efa_4 <- fa(
  corr$rho,
  nfactors = 4,
  n.obs = nrow(vic_clean),
  fm = "minres",
  rotate = "oblimin"
)
efa_4
efa_4$loadings
```

EFA4: communality and uniqueness 
```{r}
efa_4$communality
efa_4$uniquenesses
```

